Canonical rings of rational divisors on projective space and Hirzebruch surfaces

### 2015-07-11 surfaces-write-up/surface-write-up.tex
finished 1-pt generator proof, relations are left

\begin{abstract}
 	Bonding generator and relation degrees of general $\bq$-divisors
	of surfaces.
\end{abstract}

@@

\begin{rem}
Remark about how we can assume divisor only consists of hyperplanes
by taking the Veronese embedding.
\todo{Robin: do this}
\end{rem}

@@

\noindent
has a minimal presentation consisting of the $\sum_{i = 0}^{r}
{{m + i} \choose {i}}$ many generators $f_{d_i, \vec{v}} :=
\frac{u^{d_i} x^{\vec{v}}} {x_k^{c_i}}$ where $\vec{v} \in
\bz_{\geq 0}^{m + 1}$ such that $\sum_{j = 0}^{m} v_j = c_i$ and $z$ \todo{how many?}
many relations of the form
\[
	x^{\vec{u}} x^{\vec{v}} = x^{\vec{w}} x^{\vec{z}}
\]

\noindent
when $\vec{u} + \vec{v} = \vec{w} + \vec{z}$.
\end{thm}

\begin{proof}
We know $\{f_{0, \vec{u}} : \deg \vec{u} = \left\lfloor \frac{a}{b}
\right\rfloor \}$ are minimal generators in degree $1$. If
$f_{d_i, \vec{v}}$ were not minimal for $\deg \vec{v} = c_i$ for
some $i > 0$, then $f_{d_i, \vec{v}} = f_{s, \vec{w}}
f_{t, \vec{z}}$ for some $s, t \in \bn$ and $\vec{w}, \vec{z} \in
\bz_{\geq 0}^{m + 1}$. In particular, we see that 
\begin{align*}
	&d_i = s + t \\
	&c_i = \deg \vec{v} = \deg \vec{w} + \deg \vec{z}
\end{align*}

@@

Now we show that no further generators are necessary. Let $\vec{x}_i
:= (d_i, c_i) \in \bz^2$ a lattice point.

>>>>

\begin{abstract}
 	Bounding generator and relation degrees of general $\bq$-divisors
	on projective spaces $\bp^n$ and Hirzebruch surfaces.
\end{abstract}

@@

\begin{rem}
Remark about how we can assume divisor only consists of hyperplanes
by taking the Veronese embedding.
\todo{write}
\end{rem}

@@

\noindent
has a minimal presentation consisting of the $\sum_{i = 0}^{r}
{{m + i} \choose {i}}$ many generators $f_{d_i}^{\vec{v}} :=
\frac{u^{d_i} x^{\vec{v}}} {x_k^{c_i}}$ where $\vec{v} \in
\bz_{\geq 0}^{m + 1}$ such that $\sum_{j = 0}^{m} v_j = c_i$ and $z$ \todo{how many?}
many relations of the form
\[
	f_{d_i}^{\vec{v}} f_{d_j}^{\vec{w}} = f_{d_k}^{\vec{y}} f_{d_\ell}^{\vec{z}}
\]

\noindent
when $d_i + d_j = d_k + d_\ell$ and $\vec{v} + \vec{w} = \vec{y} +
\vec{z}$.
\end{thm}

\begin{proof}
We know $\{f_{0}^{\vec{v}} : \deg \vec{v} = \left\lfloor \frac{a}{b}
\right\rfloor \}$ are minimal generators in degree $1$. Now let $i >
0$. If $f_{d_i}^{\vec{v}}$ were not minimal for some $\vec{v}$ such
that $\deg \vec{v} = c_i$, then $f_{d_i}^{\vec{v}} = f_{s}^{\vec{w}}
f_{t}^{\vec{z}}$ for some $s, t \in \bn$ and $\vec{w}, \vec{z} \in
\bz_{\geq 0}^{m + 1}$. In particular, we see that 
\begin{align*}
	&d_i = s + t \\
	&c_i = \deg \vec{v} = \deg \vec{w} + \deg \vec{z}
\end{align*}

@@

Now we show that no further generators are necessary. As done in
the proof of the $\bp^1$ case by O'Dorney \cite[Theorem 6]
{dorney:canonical}, each lattice point $(d, c) \in \bz^2$ such that
$c \leq d \alpha$ is a nonnegative linear combination of $\vec{x}_i$
and $\vec{x}_{i + 1}$ where $\vec{x}_i := (d_i, c_i) \in \bz^2$.
Consequently, any element $\frac{u^{d} x^{\vec{v}}} {x_k^{c}} \in
R_D$ is expressible as
\begin{align*}
	\frac{u^{d} x^{\vec{v}}} {x_k^{c}} = \frac{u^{d_i} x^{\vec{w}}}
	{x_k^{c_i}} \frac{u^{d_{i + 1}} x^{\vec{z}}} {x_k^{c_{i + 1}}}
\end{align*}

\noindent
where $\vec{w}, \vec{z} \in \bz_{\geq 0}^{m + 1}$ such that $\deg
\vec{w} = c_i, \deg \vec{z} = c_{i + 1}$. But these are precisely
two generators of the form $f_{d_i}^{\vec{w}}$ and $f_{d_{i + 1}}
^{\vec{z}}$ as prescribed in the theorem statement.

\todo{relations}
\end{proof}
